Geometrization of characters of tori over non-Archimedean local fields

### 2013-07-18 shanghai.tex
Added the definition of bounded character sheaves.
Fundamental properties are currently under construction . . .

\newcommand{\cdef}[1]{{{\color{cyan}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\CS}{{\mathcal{C\hskip-1.8pt S}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\textit{iso}}}
\newcommand{\CSbiso}[1]{\CS_0(#1)_{/\text{iso}}}
\newcommand{\CE}{{\mathcal{C\hskip-1.8pt E}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}

@@

%We find it useful to offer an alternate perspective on a subcategory of $\CS(G)$.
In this section we describe a full subcategory of $\CS(G)$, called the category of bounded character sheaves on $G$ and denoted by $\CS_b(G)$, which will illuminate the nature of $\CS(G)$ and will also play a role in the arguments later in this paper.

In order to define $\CS_b(G)$ as must recall a fundamental fact about finite etale covers of schemes. Let $X$ be a scheme, locally of finite type over $k$, and let $\pi : Y\to X$ and $\pi_1: Y_1\to X$ be finite etale covers and let $f_1 : Y_1\to Y$ be a morphism such that $\pi_1 = \pi \circ f_1$. Then $f_1$ determines a group homomorphism $\Aut(\pi_1)\to \Aut(\pi)$, as follows  . . .

Now we may define $\CS_b(G)$, where $G$ is a smooth group scheme locally of finite type over $\Fq$.
\begin{enumerate}
\labitem{(CSb.0)}{CSb.0}
An object in $\CS_b(G)$ is a pair $(\pi,\psi)$ where $\pi : H\to G$ is a finite etale cover in the category of group schemes over $\Fq$ such that $\Gal(\bFq/\Fq)$ acts trivially on $\ker {\bar \pi}$ (so $\ker {\bar \pi} = \ker \pi$) and $\psi : \ker \pi \to \Aut(V)$ is an irreducible representation on a $\EE$-vector space $V$.
\[
\begin{tikzcd}
V & \Aut(V) & 1 \rar & \arrow[bend right,swap]{ll}{\psi} \ker \pi \rar & H \rar{\pi} & G \rar &1
\end{tikzcd}
\]
\labitem{(CSb.1)}{CSb.1}
A morphism $T : (\pi,\psi) \to (\pi',\psi')$ is a linear transformation $T : V \to V'$ which, if non-trivial, has the property that there is a finite etale cover $\pi_1 : H_1 \to G$ in the category of group schemes over $\Fq$ and morphisms $H_1\to H$, $H_1\to H'$ inducing homomorphisms $\ker \pi_1 \to \ker\pi$ and $\ker\pi_1\to \ker\pi$ such that the following diagram commutes.
\[
\begin{tikzcd}
V \arrow{dd} & \Aut(V) \ar[equal]{dd} & 1 \rar & \arrow[swap,bend left]{ll}{\psi} \ker \pi \rar & H \rar{\pi} & G \rar &1\\
 &  & 1 \rar & \ker \pi_1 \rar \uar \dar & H_1 \uar \dar \rar{\pi_1} & \ar[equal]{u} \arrow[equal]{d} G \rar &1\\
V' & \Aut(V') & 1 \rar & \arrow[bend right]{ll}{\psi'} \ker \pi' \rar & H' \rar{\pi'} & G \rar &1
\end{tikzcd}
\]
\end{enumerate}
We refer to $\CS_b(G)$ as the category of \cdef{bounded character sheaves on $G$}.
\Clifton{I'd love to find a way to define $\CS_b(G)$ without insisting that $\psi$ is an \emph{irreducible} representation, and then prove that the full subcategory of simple objects in $\CS_b(G)$ is a full subcategory of $\CS(G)$. I'm studying my notes from BIRS, letters from Masoud, and SGA7, to see how to do this.}
\begin{proposition}
$\CS_b(G)$ is a full subcategory of $\CS(G)$.
\end{proposition}

@@

>>>>

\newcommand{\cdef}[1]{{{\color{cyan}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\CS}{{\mathcal{C\hskip-1.8pt S}}}
\newcommand{\GC}{{\mathcal{G\hskip-0.8pt C}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\textit{iso}}}
\newcommand{\CSbiso}[1]{\CS_0(#1)_{/\text{iso}}}
\newcommand{\CE}{{\mathcal{C\hskip-1.8pt E}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}

@@

In this section we describe full subcategory of $\CS(G)$, 
called the category of bounded character sheaves on $G$ and denoted by $\CS_b(G)$, 
which will illuminate the nature of $\CS(G)$ and will also play a role in the arguments later in this paper.

Consider the category $\GC(G)$ of finite etale covers of $G$ that are also morphisms of abelian group schemes.
Recall that we fixed an algebraic closure $\bFq$ of $\Fq$ long ago; 
let ${\bar e} : \Spec{\bFq} \to G$ be the geometric point on $G$ obtained by composing $\Spec{\bFq} \to \Spec{\Fq}$ with the identity $e : \Spec{\Fq} \to G$ for $G$.
Let $\ker_{\bar e}$ be the set-valued functor on $\GC(G)$
defined by pull-back along the geometric point ${\bar e} : \Spec{\bFq} \to G$ 
followed by forgetting the structure sheaf and topology; 
thus, if $f : H\to G$ is a finite etale morphism of group schemes, then $\ker_{\bar e}(f)$ is the set underlying 
$\ker{\bar f} = H_{\bar e}$.
Let $\pi_1^\text{gp}(G)$ be the automorphism group of the functor $\ker_{\bar e}$. 
We will see that $\pi_1^\text{gp}(G)$ is a profinite abelian group.
A \cdef{bounded character sheaf on $G$} is a continuous irreducible representation of $\pi_1^\text{gp}(G)$. 
Let $\CS_b(G)$ be the monoidal category of irreducible continuous $\ell$-adic representations of $\pi_1^\text{gp}(G)$.

\begin{proposition}
 $G\mapsto \pi_1^\text{gp}(G)$ is a functor from group schemes over $\Fq$ to profinite abelian groups, and .....
\end{proposition}

\begin{proof}
The category $\GC(G)$, together with the set-valued functor $\ker_{\bar e}$ satisfy conditions (G1) through (G6) appearing in \cite[Exposé V, \S 4, p. 118-119]{SGA1a}.
\end{proof}

@@

\bib{SGA1a}{book}{
   author={Grothendieck, Alexander},
   title={Rev\^etements \'etales et groupe fondamental. Fasc. I: Expos\'es 1
   \`a 5},
   series={S\'eminaire de G\'eom\'etrie Alg\'ebrique},
   volume={1960/61},
   publisher={Institut des Hautes \'Etudes Scientifiques},
   place={Paris},
   date={1963},
%   pages={iv+143 pp. (not consecutively paged) (loose errata)},
%   review={\MR{0217087 (36 \#179a)}},
}
